```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(readxl)


file_path <- "fluidmilk.xlsx"
fluidmilk_data <- read_excel(file_path, sheet = "fluidmilk")
colnames(fluidmilk_data) <- c("Year", "Whole_Milk", "Reduced_Fat_2_%", 
                              "Low_Fat_1_%", "Skim", "Flavored_Whole", 
                              "Flavored_Other", "Buttermilk", "Eggnog", "Total_Sales")

fluidmilk_data <- fluidmilk_data %>%
  mutate(Year = as.numeric(Year)) %>%
  filter(!is.na(Year))

fluidmilk_data <- fluidmilk_data %>%
  mutate(across(Whole_Milk:Total_Sales, as.numeric))

#print(fluidmilk_data)

#key_stats <- fluidmilk_data %>%
#  summarise(
#    across(everything(), list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE)
#  )

#print(key_stats)
```


## Flavored Whole Milk vs Flavored Non-Whole Milk Sales Over Time

Flavored non-whole milk consistently outpaces flavored whole milk in sales, especially after the 1990s.Both categories show significant growth from the 1980s, peaking around the late 1990s to early 2000s. Sales for both categories decline sharply after 2010, aligning with broader trends of reduced fluid milk consumption.Flavored whole milk exhibits relatively stable sales compared to the larger fluctuations seen in flavored non-whole milk.


```{r,echo=FALSE,warning=FALSE,message=FALSE}

fluidmilk_data_comparison <- fluidmilk_data %>%
  mutate(
    Flavored_Whole_Category = Flavored_Whole, 
    Flavored_Non_Whole_Category = Flavored_Other 
  ) %>%
  select(Year, Flavored_Whole_Category, Flavored_Non_Whole_Category) %>%
  pivot_longer(cols = c(Flavored_Whole_Category, Flavored_Non_Whole_Category),
               names_to = "Milk_Type", values_to = "Sales")


ggplot(fluidmilk_data_comparison, aes(x = Year, y = Sales, fill = Milk_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Flavored Milk Sales: Whole vs Non-Whole", 
    x = "Year",
    y = "Sales (Million Pounds)",
    fill = "Milk Type"
  ) +
  scale_fill_manual(
    values = c("red", "gold"),
    labels = c("Flavored Whole", "Flavored Non-Whole")  
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

## T-Test Results

When camparing flavored milk vs non-whole milk sales, the t-statistic is -1.3685, which indicates the difference between the means of the two groups relative to their variability.The p-value is 0.1778, which is greater than 0.05. The mean difference between the Flavored Whole and Non-Whole ratios is -02019007 , suggesting that on average, Flavored Whole ratios are slightly lower than Non-Whole ratios. However, this difference is not significant.This means the result is not statistically significant.The confidence interval for the mean difference is [-0.049886552  0.009506417], which includes 0. This further indicates that the difference in means is not statistically significant.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
colnames(fluidmilk_data) <- c(
  "Year", "Whole", "ReducedFat2", "LowFat1", "Skim",
  "FlavoredWhole", "FlavoredOther", "Buttermilk", "Eggnog", "Total"
)

fluidmilk_clean <- fluidmilk_data %>%
  mutate(across(Whole:Total, as.numeric)) %>%
  filter(!is.na(Year)) %>%
  arrange(Year) %>% # Ensure the data is sorted by year
  mutate(NonWhole = ReducedFat2 + LowFat1 + Skim + FlavoredOther + Buttermilk + Eggnog)


fluidmilk_ratios <- fluidmilk_clean %>%
  mutate(
    FlavoredMilkRatio = FlavoredWhole / lag(FlavoredWhole),
    NonWholeRatio = NonWhole / lag(NonWhole)
  )

fluidmilk_ratios %>%
  filter(!is.na(FlavoredMilkRatio) & !is.na(NonWholeRatio)) %>%
  select(Year, FlavoredMilkRatio, NonWholeRatio) %>%
  tail(5) |>
  rename( `Flavored Milk Ratio`= FlavoredMilkRatio,
          `Non Whole Ratio` = NonWholeRatio) |>
  knitr::kable(caption= "Flavored Ratios")


#print(recent_flavored_ratios)

```


```{r,echo=FALSE,warning=FALSE,message=FALSE}

fluidmilk <- read_excel("fluidmilk.xlsx", sheet = "fluidmilk", skip = 1)

colnames(fluidmilk) <- c(
  "Year", "Whole", "ReducedFat2", "LowFat1", "Skim",
  "FlavoredWhole", "FlavoredOther", "Buttermilk", "Eggnog", "Total"
)
fluidmilk_clean <- fluidmilk %>%
  mutate(across(Whole:Total, as.numeric)) %>%
  filter(!is.na(Year)) %>%
  arrange(Year) %>% # Ensure the data is sorted by year
  mutate(NonWhole = ReducedFat2 + LowFat1 + Skim + FlavoredOther + Buttermilk + Eggnog)

sales_ratios_flavored <- fluidmilk_clean %>%
  mutate(
    FlavoredWholeRatio = FlavoredWhole / lag(FlavoredWhole),
    NonWholeRatio = NonWhole / lag(NonWhole)
  ) %>%
  select(Year, FlavoredWholeRatio, NonWholeRatio) %>%
  filter(!is.na(FlavoredWholeRatio) & !is.na(NonWholeRatio))

t.test(
  sales_ratios_flavored$FlavoredWholeRatio,
  sales_ratios_flavored$NonWholeRatio,
  paired = TRUE
) |>
  broom::tidy() |>
  knitr::kable(caption= "T-Test Results")

```
