


```{r,echo=FALSE}

library(readxl)
library(dplyr)

file_path <- "fluidmilk.xlsx"
fluidmilk_data <- read_excel(file_path, sheet = "fluidmilk")


colnames(fluidmilk_data) <- c("Year", "Whole_Milk", "Reduced_Fat_2_%", 
                              "Low_Fat_1_%", "Skim", "Flavored_Whole", 
                              "Flavored_Other", "Buttermilk", "Eggnog", "Total_Sales")


fluidmilk_data <- fluidmilk_data %>%
  mutate(Year = as.numeric(Year)) %>%
  filter(!is.na(Year))


fluidmilk_data <- fluidmilk_data %>%
  mutate(across(Whole_Milk:Total_Sales, as.numeric))

print(fluidmilk_data)

key_stats <- fluidmilk_data %>%
  summarise(
    across(everything(), list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE)
  )


print(key_stats)
```


## Flavored Whole Milk vs Flavored Non-Whole Milk Sales Over Time
```{r,echo=FALSE}
library(ggplot2)
library(tidyr)


fluidmilk_data_comparison <- fluidmilk_data %>%
  mutate(
    Flavored_Whole_Category = Flavored_Whole, 
    Flavored_Non_Whole_Category = Flavored_Other 
  ) %>%
  select(Year, Flavored_Whole_Category, Flavored_Non_Whole_Category) %>%
  pivot_longer(cols = c(Flavored_Whole_Category, Flavored_Non_Whole_Category),
               names_to = "Milk_Type", values_to = "Sales")


ggplot(fluidmilk_data_comparison, aes(x = Year, y = Sales, fill = Milk_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Flavored Whole Milk vs Flavored Non-Whole Milk Sales Over Time",
    x = "Year",
    y = "Sales (Million Pounds)",
    fill = "Milk Type"
  ) +
  scale_fill_manual(values = c("red", "gold")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )


```

Flavored non-whole milk consistently outpaces flavored whole milk in sales, especially after the 1990s.Both categories show significant growth from the 1980s, peaking around the late 1990s to early 2000s. Sales for both categories decline sharply after 2010, aligning with broader trends of reduced fluid milk consumption.Flavored whole milk exhibits relatively stable sales compared to the larger fluctuations seen in flavored non-whole milk.
