---
#title: "milkproduction"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    always_allow_html: true
---

```{r,include=FALSE,warning=FALSE,message=FALSE}
source("./scripts/load_milk_production_data.R")
source("./scripts/figure_settings.R")

milk_production = joined_data |>
  select(region,state,year,milk_production)
```
# Milk Production Statistics by State
Through analysis of the milk production dataset, we identified **California**, **Wisconsin**,  **New York**, **Pennsylvania**, **Minnesota**,  as the states with the highest milk production in the United States from 1970 to 2023. These states represent the top 5 with the largest milk production.


```{r}
top_5_states <- milk_production |>
  group_by(state) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE)) |>
  arrange(desc(total_production)) |>
  rename(State = state, 
         "Total Production" = total_production) |>
  head(5)  

total_production_usa <- round(sum(milk_production$milk_production, na.rm = TRUE))
avg_production_state <- round(mean(top_5_states$`Total Production`, na.rm = TRUE))
#min_production_state <- min(top_5_states$`Total Production`, na.rm = TRUE)

summary_stats <- tibble(
   State = c("Total Production in USA",  "Avg Production in USA"), #"Min production in USA",
  `Total Production` = c(total_production_usa, avg_production_state) #min_production_state, 
) 

final_table <- bind_rows(top_5_states, summary_stats) 
```

Additionally, total milk production in the USA over this period is approximately `r format(total_production_usa, big.mark = ",", scientific = FALSE)` million pounds,, while the average milk production is approximately `r format(avg_production_state, big.mark = ",", scientific = FALSE)` million pounds.

```{r}
#print
knitr::kable(final_table, caption = "Top 5 States and Summary Statistics for Total Production (Million Pounds) (1970-2023)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```


## Table of Top 5 States in Total Cows{.tabset}

The table below shows the top states in milk production in 1970 versus 2023. As observed in the table, the US map distribution, and the line plots, there has been a shift in regional milk production between 1970 and 2023. While Wisconsin was the leading state in 1970, California surpassed it by 2023. Additionally, states like Idaho, which played a minor role in milk production in 1970, have now become major producers.

### 1970 {#table_state_1970}

```{r}
milk_production |>
  filter(year == 1970) |>
  group_by(state) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE)) |>
  arrange(desc(total_production)) |>
  rename(State = state, 
         "Total Production 1970" = total_production) |>
  head(5)  |>
  knitr::kable(caption = "Top 5 States in Total Production (Million Pounds) (1970)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```

### 2023 {#table_state_2023}

```{r}
#table 2023 state
milk_production |>
  filter(year == 2023) |>
  group_by(state) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE)) |>
  arrange(desc(total_production)) |>
  rename(State = state, 
         "Total Production 2023" = total_production) |>
  head(5)  |>
  knitr::kable(caption = "Top 5 States in Total Production (Million Pounds) (2023)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```
## {-}
## Map of Milk Cows Distribution by State{.tabset}
### 1970{#map_state_1970}

```{r}
#merge location data with production data
us_states <- usmap::us_map()

#abbreviation for state names
data_milk_production = milk_production |> mutate(
  state_abbr = state.abb[match(state, state.name)]
)

combine_milk_production_1970 <- us_states |>
 left_join(data_milk_production, by = c("abbr" = "state_abbr")) |>
  filter (year == "1970") |>
  rename(Production = milk_production)

ggplot(combine_milk_production_1970) +
  geom_sf(aes(fill = Production), color = "white", size = 0.2) +
  #scale_fill_viridis_c(name = "milk_production",option = "plasma") +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = " Production"
  ) + 
  theme_minimal() +
  labs(title = "Milk production Distribution by State (Million Pounds) (1970)") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) -> static_map_state_1970

ggplotly(static_map_state_1970)

```

### 2023{#map_state_2023}


```{r}
#us map state 2023
combine_milk_production_2023 <- us_states |>
 left_join(data_milk_production, by = c("abbr" = "state_abbr")) |>
  filter (year == "2023") |>
  rename(Production = milk_production)

ggplot(combine_milk_production_2023) +
  geom_sf(aes(fill = Production), color = "white", size = 0.2) +
  #scale_fill_viridis_c(name = "milk_production",option = "plasma") +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Production"
  ) + 
  theme_minimal() +
  labs(title = "Milk Production Distribution by State (Million Pounds) (2023)") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
    #legend.position = "bottom"
  ) -> static_map_state_2023

ggplotly(static_map_state_2023)
```

## {-}
## Plot of Total Cows by State {.tabset}

### 1970 {#line_state_1970}

```{r}
#line plots 1970 by state
milk_production_state_1970 = milk_production |>
  filter(year == 1970) |>
  group_by(state,year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE)) |>
  arrange(desc(total_production)) |>
  rename(State = state, 
         "Total Production 1970" = total_production, 
         Year = year) 

ggplot(milk_production_state_1970, aes(x = reorder(State, `Total Production 1970`), y = `Total Production 1970`, color = State)) +
  geom_point() +  
  geom_line(group = 1, color = "lightblue") + 
  labs(
    title = "Total Production by State (Million Pounds) (1970)",
    x = "State",
    y = "Total Production",
    color = "State"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> static_line_state_1970

ggplotly(static_line_state_1970)

```
### 2023 {#lineplot2023}

```{r}
#line plot state 2023
milk_production_state_2023 = milk_production |>
  filter(year == 2023) |>
  group_by(state,year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE)) |>
  arrange(desc(total_production)) |>
  rename(State = state, 
         "Total Production 2023" = total_production, 
         Year = year)

ggplot(milk_production_state_2023, aes(x = reorder(State, `Total Production 2023`), y = `Total Production 2023`, color = State)) +
  geom_point() +  
  geom_line(group = 1, color = "lightblue") + 
  labs(
    title = "Total Production by State (Million Pounds) (2023)",
    x = "State",
    y = "Total Production",
    color = "State"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> static_line_state_2023

ggplotly(static_line_state_2023)
```

## {-}

## Plot Trend of Milk Production by State {.tabset}
### Overall {#overall_state}


```{r}
#line plot for milk production over years across states - increasing and decreasing trend
custom_colors <- c(
  "#FF9999", "#FF6666", "#FF3333", "#FF0000", "#990000", "#660000", 
  "#99FF99", "#66CC66", "#33CC33", "#009900", "#006600", "#003300", 
  "#99CCFF", "#66CCFF", "#33CCFF", "#0099FF", "#0066FF", "#0033FF", 
  "#CCCCFF", "#9999FF", "#6666FF", "#3333FF", "#0000FF", "#000099", 
  "#FF99FF", "#FF66FF", "#FF33FF", "#FF00FF", "#9900FF", "#6600FF", 
  "#6600CC", "#660099", "#660066", "#6666CC", "#9999CC", "#99CC99", 
  "#FFCC99", "#FFCC66", "#FFCC33", "#FF9966", "#FF9966", "#FF6699", 
  "#FF3366", "#FF0033", "#FF0000", "#990033", "#660033", "#330033", 
  "#9933CC", "#6600CC", "#3300CC"
)

milk_production_over_years = milk_production |>
  group_by(state, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>  
  rename(State = state, 
         Year = year, 
         `Total Production` = total_production) |> 
  ggplot(aes(x = Year, y = `Total Production`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_colour_discrete() + #don't like it
  labs(
    title = "Trend of Total Production by State (1970-2023)",
    x = "Year",
    y = "Total production",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 85, hjust = 1)
  )

ggplotly(milk_production_over_years)
```

### Increasing {#increasing_state}

```{r}
#increasing trend plot
#custom colors

custom_colors = c(
  "pink", "red2", "magenta2", "magenta3", "chartreuse3", "salmon", "maroon",
  "mediumturquoise", "purple1", "burlywood3", "orange", "blue", 
  "cyan", "yellow", "forestgreen", "dodgerblue", "gold", "orchid", 
  "plum", "black", "indianred", "violet", "darkorange", "chartreuse", "tan"
)
#increasing trend
milk_production |>
  filter(state %in% c("New Mexico","Idaho","Arizona","Colorado","Nevada","Texas",
                      "California","Washington","South Dakota","Oregon","Michigan","Utah","Kansas","Indiana","Wisconsin","Wyoming", "Georgia","New York","Pennsylvania","Vermont","Ohio","Iowa","Florida","Minnesota")) |> #
  group_by(state, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |> 
  rename(
    State = state,
    Year = year,
    `Total Production` = total_production
  ) |>
  ggplot(aes(x = Year, y = `Total Production`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_color_viridis_d(option="plasma")
  labs(
    title = "Increasing Trend of Total Production by State (1970-2023)",
    x = "Year",
    y = "Total production (Thousands)",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1)) -> state_increasing_trend
ggplotly(state_increasing_trend)

```

### Decreasing {#decreasing_state}
```{r}
#decreasing trend plot
#custom colors
custom_colors = c(
  "pink", "red2", "magenta2", "magenta3", "chartreuse3", "salmon", "maroon",
  "mediumturquoise", "purple1", "burlywood3", "orange", "blue", 
  "cyan", "yellow", "forestgreen", "dodgerblue", "gold", "orchid", 
  "plum", "black", "indianred", "violet", "darkorange", "chartreuse", "tan"
)
#decreasing trend
milk_production |>
  filter(state %in% c("Alabama", "Arkansas", "Mississippi", "Louisiana", "New Jersey",
                     "Rhode Island", "West Virginia","Tennessee", "North Dakota","Massachusetts",
                     "Missouri","South Carolina", "Delaware","Kentucky","Maryland",
                     "Oklahoma","New Hampshire", "Illinois", "North Carolina","Montana",
                     "Connecticut", "Nebraska", "Virginia", "Maine")) |>
  group_by(state, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |> 
  rename(
    Year = year, 
    State = state,
    `Total Production` = total_production
  ) |>
  ggplot(aes(x = Year, y = `Total Production`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_color_viridis_d(option="plasma")
  labs(
    title = "Decreasing Trend of Total Production by State (1970-2023)",
    x = "Year",
    y = "Total production (Thousands)",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1)
  ) -> state_decreasing_trend
ggplotly(state_decreasing_trend)

```
## {-}
## Table Trend of Total Cows by State {.tabset}
### Increasing {#increasing_state_trend_table}
```{r}
#increasing trend table
##Total production per state, year, and region
state_region_production_summary <-milk_production |>
  filter(year >= 1970 & year <= 2023) |>
  group_by(region, state, year) %>%
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop")

#year 1970
production_1970 <- state_region_production_summary |>
  filter(year == 1970) |>
  select(region, state, total_production_1970 = total_production)

#year2023
production_2023 <- state_region_production_summary |>
  filter(year == 2023) |>
  select(region, state, total_production_2023 = total_production)

#combine data
state_region_comparison <- left_join(production_1970, production_2023, by = c("region", "state")) |>
  mutate(percentage_increase = round(((total_production_2023 - total_production_1970) / total_production_1970) * 100, 2)) |>
  filter(percentage_increase > 0) |>
  arrange(desc(percentage_increase)) |>
  rename(
    Region = region,
    State = state, 
    "Total Production 1970" = total_production_1970,
    "Total Production 2023" = total_production_2023, 
    "Percentage Increase" = percentage_increase
  ) |>
  head(5)

#table
kable(state_region_comparison, caption = "States with Percentage Increase in Total Production (Million Pounds) (1970–2023)", format = "html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered") 

```
### Decreasing {#decreasing_state_trend_table}

```{r}
#decreasing trend table
state_region_production_summary <- milk_production |>
  filter(year >= 1970 & year <= 2023) |>
  group_by(region, state, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop")

#1970
production_1970 <- state_region_production_summary |>
  filter(year == 1970) |>
  select(region, state, total_production_1970 = total_production)

#2023
production_2023 <- state_region_production_summary |>
  filter(year == 2023) |>
  select(region, state, total_production_2023 = total_production)

#combine
state_region_comparison <- left_join(production_1970, production_2023, by = c("region", "state")) |>
  mutate(percentage_decrease = round(((total_production_2023 - total_production_1970) / total_production_1970) * 100, 2)) |>
  filter(percentage_decrease < 0) |>
  arrange(percentage_decrease) |>
  rename(
    Region = region,
    State = state, 
    "Total Production 1970" = total_production_1970,
    "Total Production 2023" = total_production_2023, 
    "Percentage Decrease" = percentage_decrease
  ) #|> head(5)

#table
kable(state_region_comparison, caption = "States with Percentage Decrease in Total production (Million Pounds) (1970–2023)", format = "html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")

```
## {-}

### Number of Cows by State (1970-2023)

```{r}
#us map by state 1970
#map coordinates of states
us_states <- usmap::us_map()

#abbreviation for state names
data_milk_production = milk_production |> mutate(
  state_abbr = state.abb[match(state, state.name)]
)

#merge location data with production data
combine_milk_production_1970 <- us_states |>
 left_join(data_milk_production, by = c("abbr" = "state_abbr")) |>
  filter (year == "1970") |>
  rename(" production" = milk_production)

combine_milk_production = milk_production |> 
group_by(region, state, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop")

ggplot(combine_milk_production) +
  geom_point(aes(x = reorder(state, total_production), y = total_production, color = total_production), 
             alpha = 0.7) +  
  #scale_size_continuous(name = "milk_production", range = c(3, 7)) +
 # scale_color_viridis_c(name = "milk_production")  + 
  scale_color_gradientn(
    colors = c("pink", "magenta", "orange","red"),  #magenta
    name = " Production by State"
  ) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))+
  labs(title = " Production by State (1970-2023)", 
       x = "State",
       y = " Production")

```

# Milk Production Statistics by Region

Similarly, milk production is analyzed by region. As observed in the table, the US map distribution, and the line plots, the data shows the top regions in milk production in 1970 versus 2023. In 1970, the Lake States, Appalachia, and Northeast were among the leading milk producers. By 2023, the West Coast and Mountain regions had emerged as some of the top regions in milk cow production. Overall, the data reveals significant regional shifts in milk production from 1970 to 2023, reflecting changing trends over time.

## Table of Top 5 Regions in Total Cows {.tabset}

### 1970 {#table_1970}

```{r}
#table 1970
milk_production |> 
group_by(region, year) |>
filter(year == "1970") |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_production)) |>
  select(region,total_production) |>
  rename(Region = region, 
         "Total Production 1970" = total_production) |>
  head(5)|>
  knitr::kable(caption = "Total production by Region (1970)", format = "html") |>
  kableExtra::kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```

### 2023 {#table_2023}

```{r}
#table 2023
milk_production |> 
group_by(region, year) |>
filter(year == "2023" ) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_production)) |>
  select(region, total_production) |>
  rename(Region = region, 
         "Total Production 2023" = total_production) |>
  head(5) |>
  knitr::kable(caption = "Total production by Region (2023)", format = "html") |>
  kableExtra::kable_styling(full_width = TRUE, bootstrap_options = "bordered")
  
```
## {-}

## Map of Milk Cows Distribution by Region {.tabset}

### 1970 {#map_region_1970}

```{r}
#us map 1970
us_states <- usmap::us_map()

#abbreviation for state names
data_milk_production = milk_production |> mutate(
  state_abbr = state.abb[match(state, state.name)]
)

#merge location data with production data
##region 1970
combine_region_production_1970 <- us_states |>
 left_join(data_milk_production, by = c("abbr" = "state_abbr")) |>
  filter (year == "1970") |>
  #rename(milk_production  ) |>
  group_by(region, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>
  rename(Region = region, 
         Year = year,
         `Total Production` = total_production)

ggplot(combine_region_production_1970) +
  geom_sf(aes(fill = `Total Production`), color = "white", size = 0.2) +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Production"
  )+ 
  theme_minimal() +
  labs(title = "Milk production Distribution by Region (1970)") +
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1),
    legend.position = "bottom"
  ) #-> us_plot_region_1970
#ggplotly(us_plot_region_1970)
```

### 2023 {#map_region_2023}

```{r}
#us map 2023
combine_region_production_2023 <- us_states |>
 left_join(data_milk_production, by = c("abbr" = "state_abbr")) |>
  filter (year == "2023") |>
  group_by(region, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>
  rename(Region = region, 
         Year = year,
         `Total Production` = total_production)


ggplot(combine_region_production_2023) +
  geom_sf(aes(fill = `Total Production`), color = "white", size = 0.2) +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1" 
    name = "Production"
  )+ 
  theme_minimal() +
  labs(title = "Milk production Distribution by Region (2023)") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) #-> us_plot_region_2023
#ggplotly(us_plot_region_2023)
```
## {-}

## Plot of Total Cows by Region{.tabset}

### 1970{#plot_region_1970}
```{r}
#line plot 1970
ggplot(combine_region_production_1970, aes(x = reorder(Region, `Total Production`), y = `Total Production`, color = Region)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "lightblue") +  
  scale_color_manual(values = c("Northeast" = "blue", "Lake States" = "green", 
                               "Corn Belt" = "red", "Northern Plains" = "purple", 
                               "Appalachia" = "orange", "Southeast" = "pink",
                               "Delta States" = "yellow", "Southern Plains" = "cyan", 
                               "Mountain Region" = "brown", "West Coast" = "magenta")) +
  labs(
    title = "Total production by Region (Million Pounds) (1970)",
    x = "Region",
    y = "Total Production",
    color = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> line_plot_region_2023
ggplotly(line_plot_region_2023)
```

### 2023{#plot_region_2023}

```{r}

#line plot 
ggplot(combine_region_production_2023, aes(x = reorder(Region, `Total Production`), y = `Total Production`, color = Region)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "lightblue") +  
  scale_color_manual(values = c("Northeast" = "blue", "Lake States" = "green", 
                               "Corn Belt" = "red", "Northern Plains" = "purple", 
                               "Appalachia" = "orange", "Southeast" = "pink",
                               "Delta States" = "yellow", "Southern Plains" = "cyan", 
                               "Mountain Region" = "brown", "West Coast" = "magenta")) +
  labs(
    title = "Total production by Region (Million Pounds) (1970)",
    x = "Region",
    y = "Total Production",
    color = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> line_plot_region_2023
ggplotly(line_plot_region_2023)
``` 
## {-}

## Trend of Milk Production by Region
The following plot shows the overall trend in regional milk production from 1970 to 2023. Milk production has increased over the years in most regions, including the Lake States, West Coast, and Mountain Region.

```{r}
#line plot milk production over years by region
milk_production |>
  group_by(region, year) |>
  summarise(total_production = sum(milk_production, na.rm = TRUE), .groups = "drop") |>  
  rename(
    Year = year, 
    Region = region,
    `Total Production` = total_production
  ) |>
  ggplot(aes(x = Year, y = `Total Production` , color = Region, group = Region)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  #scale_color_brewer(palette = "Set3") +  
  scale_color_manual(
    values = c("Appalachia" = "pink", 
               "Delta States" = "orange", 
               "Mountain Region" = "magenta2", 
               "Corn Belt" = "chartreuse3", 
               "Northern Plains" = "salmon", 
              "Southern Plains" = "maroon", 
              "Lake States" = "mediumturquoise",
              "Northeast" = "purple1",
              "Southeast" = "burlywood2",
                "West Coast" ="red2" )) + 
  labs(
    title = "Trend of Total Production by Region (1970-2023)",
    x = "Year",
    y = "Total production",
    color = "Region"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1),  
    legend.position = "bottom"
  ) -> trend_region
ggplotly(trend_region)
```

# Statistical Analysis by State

The results of the linear regression model for milk cow production by state and year are provided below. The state-year interaction results show how milk production (in million pounds) trends differ across states over time. For example, California exhibits substantial positive growth in milk production, with an interaction coefficient of 767.01 million pounds per year (p < 0.001), suggesting significant annual increases in production. Similarly, Texas (229.81 million pounds/year) and New Mexico (208.66 million pounds/year) also show notable increases over time. In contrast, states like Tennessee (-23.16 million pounds/year) experience declines in milk production, indicating a regional decrease in dairy production. The year variable itself has a negative coefficient of -15.24 million pounds (p = 0.009), suggesting an overall decline in milk production across all states over time, though this trend is not uniform.

```{r}
#lm model milk production by state
milk_production |> 
  mutate(
    state = as.factor(state),  
    year = as.numeric(year)     
  ) |> 
  lm(milk_production ~ state * year, data = _) |>  
  broom::tidy() |>
  head(5) |>
  knitr::kable(digits=3, format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")

```

# Statistical Analysis by Region

The linear regression model results for milk production by region and year, with Appalachia as the reference region, are given below. While the year variable itself has a non-significant effect on milk production across all regions (estimate = -20.576, p = 0.237), significant region-year interactions indicate that some regions have seen increases in milk production over time. Notably, the Lake States (estimate = 130.39, p < 0.001), Mountain Region (estimate = 114.22, p < 0.001), and West Coast (estimate = 314.34, p < 0.001) have experienced positive trends in milk production over time, showing significant growth.

```{r}
#lm model milk production by region 
milk_production |> 
  mutate(
    #region = as.factor(region), 
    #region = relevel(as.factor(region), ref = "Southern Plains"),
    year = as.numeric(year)     
  ) |> 
  lm(milk_production ~ region * year, data = _) |>  
  broom::tidy() |>
  #head(5) |>
  knitr::kable(digits=3, format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```

