---
#title: "Milk Cows per Thousand Head Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    always_allow_html: true
---

```{r, include=FALSE, warning=FALSE,message=FALSE}
source("./scripts/load_milk_production_data.R")
source("./scripts/figure_settings.R")

milk_cows = joined_data |>
  select(region,state,year,number_of_cows)

```

# Milk Cows Statistics by State

Through analysis of the milk cows dataset, we identified **Wisconsin**, **California**, **New York**, **Minnesota**, and **Pennsylvania** as the states with the highest total number of milk cows in the United States from 1970 to 2023. These states represent the top 5 with the largest number of cows, as determined by summing the total cows per state.

```{r}
top_5_states <- milk_cows |>
  group_by(state) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE)) |>
  arrange(desc(total_cows)) |>
  rename(State = state, 
         "Total Cows" = total_cows) |>
  head(5)  

total_cows_usa <- round(sum(milk_cows$number_of_cows, na.rm = TRUE))
avg_cows_state <- round(mean(top_5_states$`Total Cows`, na.rm = TRUE))
#min_cows_state <- min(top_5_states$`Total Cows`, na.rm = TRUE)

summary_stats <- tibble(
   State = c("Total Cows in USA",  "Average Number of Cows in USA"), #"Min Number of Cows in USA",
  `Total Cows` = c(total_cows_usa, avg_cows_state) #min_cows_state, 
) 

final_table <- bind_rows(top_5_states, summary_stats) 
```
Additionally, the total number of cows in the USA over this period is approximately `r format(total_cows_usa, big.mark = ",", scientific = FALSE)`, while the average number of cows in the top 5 states is approximately `r format(avg_cows_state, big.mark = ",", scientific = FALSE)`.
```{r}
#print
knitr::kable(final_table, caption = "Top 5 States and Summary Statistics for Total Cows (per Thousand Head) (1970-2023)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```
The table below shows the top states in total milk cows in 1970 vs 2023. As observed in the table, the US map distribution, and the line plots, there has been a shift in regional milk cow production between 1970 and 2023. While Wisconsin was the leading state in 1970, California surpassed it by 2023. Additionally, states like Idaho, which did not play a major role in milk production in 1970, have now become one of the leading producers.

## Table of Top 5 States in Total Cows{.tabset}
### 1970 {#table_state_1970}
```{r}
#table 1970 by state
milk_cows |>
  filter(year == 1970) |>
  group_by(state) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE)) |>
  arrange(desc(total_cows)) |>
  rename(State = state, 
         "Total Cows 1970" = total_cows) |>
  head(5)  |>
  knitr::kable(caption = "Top 5 States in Total Cows (per Thousand Head) (1970)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```

### 2023 {#table_state_2023}
```{r}
#table 2023 by state
milk_cows |>
  filter(year == 2023) |>
  group_by(state) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE)) |>
  arrange(desc(total_cows)) |>
  rename(State = state, 
         "Total Cows 2023" = total_cows) |>
  head(5)  |>
  knitr::kable(caption = "Top 5 States in Total Cows (per Thousand Head) (2023)", format="html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")
```
## {-}
## Map of Milk Cows Distribution by State{.tabset}
### 1970{#map_state_1970}
```{r}
#us map by state 1970
#map coordinates of states
us_states <- usmap::us_map()

#abbreviation for state names
data_milk_cows = milk_cows |> mutate(
  state_abbr = state.abb[match(state, state.name)]
)

#merge location data with production data
combine_milk_cows_1970 <- us_states |>
 left_join(data_milk_cows, by = c("abbr" = "state_abbr")) |>
  filter (year == "1970") |>
  rename("Number of Cows" = number_of_cows)

ggplot(combine_milk_cows_1970) +
  geom_sf(aes(fill = `Number of Cows`), color = "white", size = 0.2) +
  #scale_fill_viridis_c(name = "number_of_cows",option = "plasma") +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Number of Cows"
  ) + 
  theme_minimal() +
  labs(title = "Milk Cows Distribution by State (per Thousand Head) (1970)") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) -> static_map_state_1970

ggplotly(static_map_state_1970)
```
### 2023{#map_state_2023}
```{r}
#map coordinates of states
us_states <- usmap::us_map()

#abbreviation for state names
data_milk_cows = milk_cows |> mutate(
  state_abbr = state.abb[match(state, state.name)]
)

#us map state 2023
combine_milk_cows_2023 <- us_states |>
 left_join(data_milk_cows, by = c("abbr" = "state_abbr")) |>
  filter (year == "2023") |>
  rename("Number of Cows" = number_of_cows)

ggplot(combine_milk_cows_2023) +
  geom_sf(aes(fill = `Number of Cows`), color = "white", size = 0.2) +
  #scale_fill_viridis_c(name = "number_of_cows",option = "plasma") +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Number of Cows"
  ) + 
  theme_minimal() +
  labs(title = "Milk Cows Distribution by State (per Thousand Head) (2023)") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
    #legend.position = "bottom"
  ) -> static_map_state_2023

ggplotly(static_map_state_2023)
```

## {-}
## Plot of Total Cows by State {.tabset}
### 1970 {#line_state_1970}
```{r}
#line plots 1970 by state
milk_cows_state_1970 = milk_cows |>
  filter(year == 1970) |>
  group_by(state,year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE)) |>
  arrange(desc(total_cows)) |>
  rename(State = state, 
         "Total Cows 1970" = total_cows, 
         Year = year) 
ggplot(milk_cows_state_1970, aes(x = reorder(State, `Total Cows 1970`), y = `Total Cows 1970`, color = State)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "lightblue") + 
  labs(
    title = "Total Number of Cows by State (per Thousand Head) (1970)",
    x = "State",
    y = "Total Cows",
    color = "State"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> static_line_state_1970

ggplotly(static_line_state_1970)
```
### 2023 {#lineplot2023}
```{r}
#line plot state 2023
milk_cows_state_2023 = milk_cows |>
  filter(year == 2023) |>
  group_by(state,year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE)) |>
  arrange(desc(total_cows)) |>
  rename(State = state, 
         "Total Cows 2023" = total_cows, 
         Year = year)
ggplot(milk_cows_state_2023, aes(x = reorder(State, `Total Cows 2023`), y = `Total Cows 2023`, color = State)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "lightblue") + 
  labs(
    title = "Total Number of Cows by State (per Thousand Head) (2023)",
    x = "State",
    y = "Total Cows",
    color = "State"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> static_line_state_2023

ggplotly(static_line_state_2023)

```
## {-}
The plots and tables below illustrate the overall trend in the total number of milk cows from 1970 to 2023. The data shows a shift in production, with some states experiencing a significant increase in milk cow numbers, while others have seen a decline. States like New Mexico, Idaho, Arizona, and Colorado have experienced more than a 100% increase in milk cow production. In contrast, states like Alabama, Mississippi, Louisiana, and New Jersey have seen a near-total decline in their milk cow industries.

## Plot Trend of Total Cows by State {.tabset}
### Overall {#overall_state}
```{r}
#line plot for milk cows over years across states - increasing and decreasing trend
custom_colors <- c(
  "#FF9999", "#FF6666", "#FF3333", "#FF0000", "#990000", "#660000", 
  "#99FF99", "#66CC66", "#33CC33", "#009900", "#006600", "#003300", 
  "#99CCFF", "#66CCFF", "#33CCFF", "#0099FF", "#0066FF", "#0033FF", 
  "#CCCCFF", "#9999FF", "#6666FF", "#3333FF", "#0000FF", "#000099", 
  "#FF99FF", "#FF66FF", "#FF33FF", "#FF00FF", "#9900FF", "#6600FF", 
  "#6600CC", "#660099", "#660066", "#6666CC", "#9999CC", "#99CC99", 
  "#FFCC99", "#FFCC66", "#FFCC33", "#FF9966", "#FF9966", "#FF6699", 
  "#FF3366", "#FF0033", "#FF0000", "#990033", "#660033", "#330033", 
  "#9933CC", "#6600CC", "#3300CC"
)

milk_cows_over_years = milk_cows |>
  group_by(state, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>  
  rename(State = state, 
         Year = year, 
         `Total Cows` = total_cows) |> 
  ggplot(aes(x = Year, y = `Total Cows`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_colour_discrete() + #don't like it
  labs(
    title = "Trend of Total Cows by State (1970-2023)",
    x = "Year",
    y = "Total Number of Cows",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 85, hjust = 1)
  )

ggplotly(milk_cows_over_years)
```
### Increasing {#increasing_state}
```{r}
#increasing trend plot
#custom colors
custom_colors = c("pink","red2","magenta2","magenta3","chartreuse3","salmon","maroon",
                  "mediumturquoise","purple1","burlywood3","orange","blue")

#increasing trend
milk_cows |>
  filter(state %in% c("New Mexico","Idaho","Arizona","Colorado","Nevada",
                      "California","Texas","Washington","Oregon","Utah","South Dakota","Michigan")) |>
  group_by(state, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |> 
  rename(
    State = state,
    Year = year,
    `Total Cows` = total_cows
  ) |>
  ggplot(aes(x = Year, y = `Total Cows`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_color_viridis_d(option="plasma")
  labs(
    title = "Increasing Trend of Total Cows by State (1970-2023)",
    x = "Year",
    y = "Total Number of Cows (Thousands)",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1)) -> state_increasing_trend
ggplotly(state_increasing_trend)

```
### Decreasing {#decreasing_state}
```{r}
#decreasing trend plot
#custom colors
custom_colors <- c(
  "#FF9999", "#FF6666", "#FF3333", "#FF0000", "#990000", "#660000", 
  "#99FF99", "#66CC66", "#33CC33", "#009900", "#006600", "#003300", 
  "#99CCFF", "#66CCFF", "#33CCFF", "#0099FF", "#0066FF", "#0033FF", 
  "#CCCCFF", "#9999FF", "#6666FF", "#3333FF", "#0000FF", "#000099", 
  "#FF99FF", "#FF66FF", "#FF33FF", "#FF00FF", "#9900FF", "#6600FF", 
  "#6600CC", "#660099", "#660066", "#6666CC", "turquoise", "#99CC99", 
  "#FFCC99", "#FFCC66", "#FFCC33", "#FF9966", "#FF9966", "#FF6699", 
  "#FF3366", "#FF0033", "#FF0000", "#990033", "#660033", "#330033", 
  "#9933CC", "#6600CC", "#3300CC"
)

#decreasing trend
milk_cows |>
  filter(state %in% c("Alabama", "Arkansas", "Mississippi", "Louisiana", "New Jersey",
                     "Rhode Island", "Tennessee", "North Dakota", "West Virginia",
                     "South Carolina", "Kentucky", "Massachusetts", "Missouri",
                     "Delaware", "North Carolina", "Montana", "Maryland", "Illinois", "Oklahoma",
                     "New Hampshire", "Connecticut", "Nebraska", "Virginia", "Maine",
                     "Minnesota", "Iowa", "Florida", "Wyoming", "Ohio", "Vermont", "Georgia",
                     "Pennsylvania", "New York", "Wisconsin", "Indiana", "Kansas")) |>
  group_by(state, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>  
  rename(
    Year = year, 
    State = state,
    `Total Cows` = total_cows
  ) |>
  ggplot(aes(x = Year, y = `Total Cows`, color = State, group = State)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  scale_color_manual(values = custom_colors) +  
  #scale_color_viridis_d(option="plasma")
  labs(
    title = "Decreasing Trend of Total Cows by State (1970-2023)",
    x = "Year",
    y = "Total Number of Cows (Thousands)",
    color = "State"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1)
  ) -> state_decreasing_trend
ggplotly(state_decreasing_trend)

```

## {-}
## Table Trend of Total Cows by State {.tabset}
### Increasing {#increasing_state_trend_table}
```{r}
#increasing trend table
##total number of cows per state, year, and region
state_region_cows_summary <-milk_cows |>
  filter(year >= 1970 & year <= 2023) |>
  group_by(region, state, year) %>%
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop")

#year 1970
cows_1970 <- state_region_cows_summary |>
  filter(year == 1970) |>
  select(region, state, total_cows_1970 = total_cows)

#year2023
cows_2023 <- state_region_cows_summary |>
  filter(year == 2023) |>
  select(region, state, total_cows_2023 = total_cows)

#combine data
state_region_comparison <- left_join(cows_1970, cows_2023, by = c("region", "state")) |>
  mutate(percentage_increase = round(((total_cows_2023 - total_cows_1970) / total_cows_1970) * 100, 2)) |>
  filter(percentage_increase > 0) |>
  arrange(desc(percentage_increase)) |>
  rename(
    Region = region,
    State = state, 
    "Total Cows 1970" = total_cows_1970,
    "Total Cows 2023" = total_cows_2023, 
    "Percentage Increase" = percentage_increase
  )|>
  head(5)

#table
kable(state_region_comparison, caption = "States with Percentage Increase in Total Number of Cows (Per Thousand Head) (1970–2023)", format = "html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered") 

```
### Decreasing {#decreasing_state_trend_table}
```{r}
#decreasing trend table
state_region_cows_summary <- milk_cows |>
  filter(year >= 1970 & year <= 2023) |>
  group_by(region, state, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop")

#1970
cows_1970 <- state_region_cows_summary |>
  filter(year == 1970) |>
  select(region, state, total_cows_1970 = total_cows)

#2023
cows_2023 <- state_region_cows_summary |>
  filter(year == 2023) |>
  select(region, state, total_cows_2023 = total_cows)

#combine
state_region_comparison <- left_join(cows_1970, cows_2023, by = c("region", "state")) |>
  mutate(percentage_decrease = round(((total_cows_2023 - total_cows_1970) / total_cows_1970) * 100, 2)) |>
  filter(percentage_decrease < 0) |>
  arrange(percentage_decrease) |>
  rename(
    Region = region,
    State = state, 
    "Total Cows 1970" = total_cows_1970,
    "Total Cows 2023" = total_cows_2023, 
    "Percentage Decrease" = percentage_decrease
  ) |>
  head(5)

#table
kable(state_region_comparison, caption = "States with Percentage Decrease in Total Number of Cows (Per Thousand Head) (1970–2023)", format = "html") |>
  kable_styling(full_width = TRUE, bootstrap_options = "bordered")

```
## {-}


### Number of Cows by State (1970-2023)
```{r}
combine_milk_cows = milk_cows |> 
group_by(region, state, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop")

ggplot(combine_milk_cows) +
  geom_point(aes(x = reorder(state, total_cows), y = total_cows, color = total_cows), 
             alpha = 0.7) +  
  #scale_size_continuous(name = "number_of_cows", range = c(3, 7)) +
 # scale_color_viridis_c(name = "number_of_cows")  + 
  scale_color_gradientn(
    colors = c("pink", "magenta", "orange","red"),  #magenta
    name = "Number of Cows by State"
  ) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))+
  labs(x = "State", #title = "Number of Cows by State (1970-2023)", 
       y = "Number of Cows")

```

# Milk Cows Statistics by Region

Similarly, milk cow production is analyzed by region. As observed in the table, the US map distribution, and the line plots, the data shows the top regions in total milk cows in 1970 vs 2023. In 1970, the Lake States, with 3,196 thousand milk cows, were the leading region, followed by other key regions such as the Northeast, Corn Belt, Appalachia, and West Coast. By 2023, the Mountain region had emerged as one of the top regions in milk cow production. Overall, the data reveals significant regional shifts in milk cow production from 1970 to 2023, reflecting changing trends over time.

## Table of Top 5 Regions in Total Cows {.tabset}

### 1970 {#table_1970}
```{r}
#region 1970
combine_region_cows_1970 <- us_states |>
 left_join(data_milk_cows, by = c("abbr" = "state_abbr")) |>
  filter (year == "1970") |>
  group_by(region, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>
  rename(Region = region, 
         Year = year,
         `Total Cows` = total_cows)

#table 1970
milk_cows |> 
group_by(region, year) |>
filter(year == "1970") |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_cows)) |>
  select(region,total_cows) |>
  rename(Region = region, 
         "Total Cows 1970" = total_cows) |>
  head(5) |>
  knitr::kable(caption = "Total Number of Cows by Region (1970)", format = "html") |>
  kableExtra::kable_styling(full_width = TRUE, bootstrap_options = "bordered")


```
### 2023 {#table_2023}
```{r}
#region 2023
combine_region_cows_2023 <- us_states |>
 left_join(data_milk_cows, by = c("abbr" = "state_abbr")) |>
  filter (year == "2023") |>
  group_by(region, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>
  rename(Region = region, 
         Year = year,
         `Total Cows` = total_cows)


#table 
milk_cows |> 
group_by(region, year) |>
filter(year == "2023" ) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_cows)) |>
  select(region, total_cows) |>
  rename(Region = region, 
         "Total Cows 2023" = total_cows) |>
  head(5)|>
  knitr::kable(caption = "Total Number of Cows by Region (2023)", format = "html") |>
  kableExtra::kable_styling(full_width = TRUE, bootstrap_options = "bordered")
  
```
## {-}

## Map of Milk Cows Distribution by Region {.tabset}

### 1970 {#map_region_1970}

```{r}
#us map 1970
ggplot(combine_region_cows_1970) +
  geom_sf(aes(fill = `Total Cows`), color = "white", size = 0.2) +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Number of Cows"
  )+ 
  theme_minimal() +
  labs(title = "Milk Cows Distribution by Region (1970)") +
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1),
    legend.position = "bottom"
  ) #-> us_plot_region_1970
#ggplotly(us_plot_region_1970)
```

### 2023 {#map_region_2023}

```{r}
#us map 2023
ggplot(combine_region_cows_2023) +
  geom_sf(aes(fill = `Total Cows`), color = "white", size = 0.2) +
  scale_fill_gradientn(
    #colors = c("pink", "magenta","orange","red"),  #magenta
    colors = c("lavender","magenta","darkmagenta","red2"), #"purple1"
    name = "Number of Cows"
  )+ 
  theme_minimal() +
  labs(title = "Milk Cows Distribution by Region (2023)") +
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1),
    legend.position = "bottom"
  ) #-> us_plot_region_2023
#ggplotly(us_plot_region_2023)

```
## {-}

## Plot of Total Cows by Region{.tabset}

### 1970{#plot_region_1970}
```{r}
#line plot 1970
ggplot(combine_region_cows_1970, aes(x = reorder(Region, `Total Cows`), y = `Total Cows`, color = Region)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "gray") +  
  scale_color_manual(values = c("Northeast" = "blue", "Lake States" = "green", 
                               "Corn Belt" = "red", "Northern Plains" = "purple", 
                               "Appalachia" = "orange", "Southeast" = "pink",
                               "Delta States" = "yellow", "Southern Plains" = "cyan", 
                               "Mountain Region" = "brown", "West Coast" = "magenta")) +
  labs(
    title = "Total Number of Cows by Region (per Thousand Head) (1970)",
    x = "Region",
    y = "Total Cows",
    color = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> line_plot_region_1970
ggplotly(line_plot_region_1970)


```

### 2023{#plot_region_2023}

```{r}
#line plot 
ggplot(combine_region_cows_2023, aes(x = reorder(Region, `Total Cows`), y = `Total Cows`, color = Region)) +
  geom_point(size=0.8) +  
  geom_line(group = 1, color = "gray") +  
  scale_color_manual(values = c("Northeast" = "blue", "Lake States" = "green", 
                               "Corn Belt" = "red", "Northern Plains" = "purple", 
                               "Appalachia" = "orange", "Southeast" = "pink",
                               "Delta States" = "yellow", "Southern Plains" = "cyan", 
                               "Mountain Region" = "brown", "West Coast" = "magenta")) +
  labs(
    title = "Total Number of Cows by Region (per Thousand Head) (1970)",
    x = "Region",
    y = "Total Cows",
    color = "Region"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) -> line_plot_region_2023
ggplotly(line_plot_region_2023)

```
## Trend of Total Cows by Region (1970-2023)
The following plot shows the overall trend in regional milk cow production from 1970 to 2023. Milk cow production in the Lake States, Northeast, Corn Belt, and Appalachia has decreased over time. In contrast, milk cow production in the West Coast, Mountain Region, and Southern Plains has increased, highlighting the regional shifts in milk cow production over the past few decades.
```{r}
#line plot milk cows over years by region
milk_cows |>
  group_by(region, year) |>
  summarise(total_cows = sum(number_of_cows, na.rm = TRUE), .groups = "drop") |>  
  rename(
    Year = year, 
    Region = region,
    `Total Cows` = total_cows
  ) |>
  ggplot(aes(x = Year, y = `Total Cows` , color = Region, group = Region)) +  
  geom_point(size=0.8) +  
  geom_line() +    
  #scale_color_brewer(palette = "Set3") +  
  scale_color_manual(
    values = c("Appalachia" = "pink", 
               "Delta States" = "orange", 
               "Mountain Region" = "magenta2", 
               "Corn Belt" = "chartreuse3", 
               "Northern Plains" = "salmon", 
              "Southern Plains" = "maroon", 
              "Lake States" = "mediumturquoise",
              "Northeast" = "purple1",
              "Southeast" = "burlywood2",
                "West Coast" ="red2" )) + 
  labs(
    #title = "Trend of Total Cows by Region (1970-2023)",
    x = "Year",
    y = "Total Number of Cows",
    color = "Region"
  ) +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 80, hjust = 1),  
    legend.position = "bottom"
  ) -> trend_region
ggplotly(trend_region)
```

# Statistical Analysis by State
The linear regression model for milk cow production by state and year results are given below. States like California (-50,064.73, p < 0.001), Minnesota (19,388.98, p < 0.001), and Wisconsin (27,158.24, p < 0.001) have significant positive deviations in milk cow populations compared to the reference state (Alabama), indicating large milk cow populations. The year variable has a significant negative effect on milk cow numbers (-1.983, p < 0.001), suggesting a general decline in milk cow production across all states over time. Significant State-Year interactions (e.g., California:year 25.73, p < 0.001; New Mexico:year 9.48, p < 0.001) suggest that some states, such as California, have seen increases in milk cow production over time, while others have seen declines.


```{r}
#lm model milk production by state
milk_cows |> 
  mutate(
    state = as.factor(state),  
    year = as.numeric(year)     
  ) |> 
  lm(number_of_cows ~ state * year, data = _) |>  
  broom::tidy() |>
  head(5) |>
  knitr::kable(digits = 3)


#milk_cows |> 
#  mutate(
#    region = as.factor(state), 
#    year = as.numeric(year)) |>
#  aov(number_of_cows ~ state*year, data = _) |>  
#  broom::tidy() |>
#  knitr::kable(digits=3)

```
# Statistical Analysis by Region

The linear regression model for milk cow production by region and year reveals that California (-23,512.16, p < 0.001) and the Mountain Region (-12,700.40, p < 0.001) show significant negative deviations in milk cow populations compared to the reference region (Appalachia), though California saw an increase in production over time (positive interaction term with year: 12.00, p < 0.001), while Appalachia experienced a decline. The year variable has a significant negative effect on milk cow numbers (-3.24, p < 0.001), indicating a general decline in milk cow populations across all regions over time. Significant State-Year interactions (e.g., Mountain Region:year 6.36, p < 0.001; West Coast:year 12.00, p < 0.001) suggest that some regions, like California, have seen increases in milk cow production, while regions like Appalachia have seen declines.

```{r}
#lm model milk production by region 
milk_cows |> 
  mutate(
    region = as.factor(region),  
    year = as.numeric(year)     
  ) |> 
  lm(number_of_cows ~ region * year, data = _) |>  
  broom::tidy() |>
  head(5) |>
  knitr::kable(digits=3)
```